Astronomical Image colorization and up-scaling with generative adversarial networks\\

### 2021-09-23 Research Paper/paper.tex
Added figures
Added mathematical model

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%\usepackage{cite}
\usepackage[authoryear]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{hyperref} 
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

As noted in \citep{Gao2019astronomical}, a huge number of raw images are present unprocessed and unnoticed in the Hubble Legacy Archives. These raw images are typically black and white, low-resolution and unfit to be shared with the world. It takes huge amounts of hours to process them. This processing is necessary because it's difficult for astronomers to distinguish objects from the raw images. The processing is further made necessary owing to noise from other bodies in the universe, changing optical characteristics in the system and random \& synthetic noise from the sensors in the telescope. Furthermore, for the process of highlighting small features that ordinarily wouldn't be able to be picked out against noise of the image, we need colorization. The processing of the images is so time consuming that the images are rarely seen by human eyes. Not only is new data being continuously produced by Hubble Telescope, but new telescopes are soon to come online. This goes to show that the problem will likely get worse. A simplification of image processing by using artificial image colorization and super-resolution can be done in an automated fashion to make it easier for astronomers to visually identify and analyze objects in Hubble dataset.

@@

We train models using both color spaces and compare their performance.
\subsection{Transferred learning and model tweaking}
\hspace*{0.25 in}\cite{isola2018imagetoimage} proposed a general solution to many image-to-image translation tasks. We propose a similar methodology as proposed in the paper with a few minor tweaks, so as to significantly reduce the amount of training data needed and minimize the training time to achieve similar results. Initially, we propose the use of a U-net for the generator. A pre-trained ResNet-18 network was used as the \textit{encoder} for the U-net with half of it's layers clipped off so as to get feature abstractions of the intermediate layers. \cite{ledig2017photorealistic} showed that training the generator separately in a supervised, deterministic manner helps it generalize the mapping from the input space to outputs. The idea solves the problem of \textit{"The blind leading the blind"} that is persistent in most image-to-image translation tasks to date. \\
\hspace*{0.25 in}We pre-train our generator independently using the imagenet weights over the Common Objects in Context (COCO) dataset. Unlike adversarial training, this phase was supervised and the loss function used was \textit{mean absolute error} or the \textit{L1 Norm} from the target image. Though trained deterministically,the problem of rectifying incorrect predictions still persists due to constraints over the convergence of loss metric. To combat this, this trained generator was trained once again in an adversarial fashion to help generalize it further. We hypothesize that re-training in an adversarial fashion will further rectify the subtle color differences that \textit{mae} couldn't solve.\\
\hspace*{0.25 in}A pre-trained ResNet-50 with imagenet weights was used as the discriminator with last few layers clipped off. The discriminative network used is something called a "Patch Discriminator" proposed by \cite{isola2018imagetoimage}. In a \textit{vanilla} discriminator proposed by \cite{goodfellow2014generative}, the network produces a scalar output, representing the probability that the data $x$ belongs to the input distribution and not the generator distribution $p_g$. \cite{isola2018imagetoimage} proposed a modification to the discriminative network so as to produce, instead of a single scalar, a vector of probabilities representing different localities of the input distribution (image) $x$. For instance, a discriminator producing a $50 \times 50$ vector represents the probabilities every receptive field that is covered by the output vector. Thus, we can localize the corrections in the image that the generator should make.\\
\hspace*{0.25 in}Finally, the trained generator and trained discriminator are fine-tune to fit on our data which is rather small in size compared to the previous datasets. These networks are trained in an adversarial fashion using the conditional GAN objective function \citep{isola2018imagetoimage} with some noise introduced as the L1 norm of generated image tensor to the target image tensor. The reason behind this is to train the generator using an adversarial component while trying to minimize the Manhattan distance between the generator output and target vector space.
\subsection{Experimental Setup}
\hspace*{0.25 in}We use a mini-batch gradient descent with a batch size of 10, 16 and 32 for different iterations. We use Adam optimizer with $\beta_1 = 0.5$ and $\beta_2 = 0.999$. The generator and discriminator has a learning rate of $2e-4$ which remains constant throughout the training. We train the model with different epochs ranging between 20,50 and 100, saving the best model weights determined by L1 norm between the output of the generator and the target image. We use early stopping with a patience value of 10. The image size is $64\times64$. Our implementation is using Python, numpy, Tensorflow and tf-keras. It takes about 24 hours to complete training over the COCO dataset and about 12 to 13 hours to fit the model on given astronomical data on a NVIDIA Tesla K80 GPU.

\section{Results and Discussions}
\hspace*{0.25 in}In the following section, we briefly compare and summarize the results of the implemented architectures and evaluate their performance. The evaluation is done qualitatively as well as quantitatively. We compare the performance of each model using L1 and L2 norm distance between the predictions and targets. Though unreliable, this method provides us with a somewhat decent ground to perform a comparative study and evaluate the reliability of such metrics on GAN evaluation compared to qualitative, visual evaluation. We also measure the performance of the GANs by using Fréchet inception distance (FID) as an attempt to quantify how close the produced images are to the original ones.\\
\hspace*{0.25 in}To evaluate the model performance by virtue of convergence of the objective function, we plot the generator loss throughout the training on two different networks.\\
\begin{figure}[!htb]
	\centering
	\begin{subfigure}[b]{0.25\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/resnet18_generator_loss}
		\caption{ResNet18 U-net Generator}
		\label{fig: resnet_loss}
	\end{subfigure}
		\hspace{0.1 in}
	\begin{subfigure}[b]{0.25\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/resnet18_discriminator_loss}
		\caption{ResNet50 discriminator loss}
		\label{fig: resnet50_disc_loss}
	\end{subfigure}
	\caption{Curve plots for ResNet18 U-net \& ResNet50 discriminator}
	\label{fig: resnet18_curves}
\end{figure}
Figure \ref{fig: resnet_loss} and \ref{fig: resnet50_disc_loss} shows that the generator converges quite nicely. The discriminator on the other hand oscillates because of the convergence that the generator shows. GAN losses are pretty non-intuitive but we can draw some observations from the loss curves. It seems that the pattern of oscillation and convergence repeats when networks are trained in an adversarial fashion. It is observed that the ResNet U-net with a pre-trained ResNet-18 for its backbone converges decently in the beginning but later shows spikes when nearing the end of training loop. It seems that the loss convergence doesn't signify whether the model is predicting expected results. The loss function just converges to the minimum value that the cost function descends to, permitted by the learning rate. It would normally signify that the GAN has found some optimum point in the vector space that is at the lowest potential and cant decrease any further. It essentially means that the GAN has leaned enough. Due to the large number of dimensions, owing to the high amount of trainable variables, such combinations, where the function converges, can be huge in volume. Thus, these numbers don't provide any better understanding of the bias or variance the model is facing. Also, we discover that if the loss hasn't converged well, it doesn't necessarily mean that the model hasn't learned anything. On visual inspection, the generated results show similar results to the ground truth, even with high generator losses. This might be due to presence of a content loss parameter in the loss function where we try to minimize the function by  minimizing the L1 norm between the generator predictions and target.

@@

\pagebreak
Table \ref{tab:Super-resolution_results} shows the L1 and L2 distances of predicted results by each model with the ground truth image. It is observed that Ledig's SRGAN, after a bit of fine-tuning performs really well in comparison to the the pre-trained version. To further improve this, \cite{lim2017enhanced} proposed an optimized version of SRGAN by removing the unnecessary modules in the conventional resnets and showed that Enhanced Deep Residual Networks performed better at upscaling task. When trained in an adversarial manner, the ED-SRGAN performs better than the traditional SRGAN. \cite{yu2018wide} further improved the idea by increasing the widening factor ($\times 2$ and $\times 4$) to ($\times 6$ and $\times 9$). This Wide Activation Deep Super Resolution network further improved the performance for single image super-resolution. When we implement this in an adversarial manner, we achieve excellent results. It is evident from the results that the best performing network is the WDSR.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%\usepackage{cite}
\usepackage[authoryear]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{hyperref} 
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}} 
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{varwidth}
\usepackage{tikz}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\usetikzlibrary{chains, calc,fit}

@@

As noted in \citep{Gao2019astronomical}, a huge number of raw images are present unprocessed and unnoticed in the Hubble Legacy Archives. These raw images are typically black and white, low-resolution and unfit to be shared with the world. It takes huge amounts of hours to process them. This processing is necessary because it's difficult for astronomers to distinguish objects from the raw images. The processing is further made necessary owing to noise from other bodies in the universe, changing optical characteristics in the system and random \& synthetic noise from the sensors in the telescope. Furthermore, for the process of highlighting small features that ordinarily wouldn't be able to be picked out against noise of the image, we need colorization. The processing of the images is so time consuming that the images are rarely seen by human eyes. Not only is new data being continuously produced by Hubble Telescope, but new telescopes are soon to come online. This goes to show that the problem will likely get worse. A simplification of image processing by using artificial image colorization and super-resolution can be done in an automated fashion to make it easier for astronomers to visually identify and analyze objects in Hubble dataset.
\begin{figure}[!htb]
	\centering
	\begin{subfigure}[b]{0.2\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/original_sample}
		\caption{Original image}
		\label{fig: original_sample}
	\end{subfigure}
	\begin{subfigure}[b]{0.2\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/GAN_output_sample}
		\caption{GAN output}
		\label{fig: gan_output_sample}
	\end{subfigure}
	\caption{Fig \ref{fig: original_sample} is the original image and fig \ref{fig: gan_output_sample} is the image produced by our GAN}
	\label{fig: sample_comp}
\end{figure}

@@

We train models using both color spaces and compare their performance.
\subsection{Mathematical Model}
\subsubsection{Generative Adversarial Networks}
 A generative network, $G$, is supposed to learn the underlying distribution of a latent space, Y. Instead of visually assessing the quality of network outputs and judge how we can adapt the network to produce convincing results, we incorporate automatic tweaking during training by introducing a discriminative network $D$. The network D takes in both the fabricated outputs generated by G and real inputs from the underlying distribution Y. The network produces a probability of the image belonging to the real or fabricated space.
\begin{figure*}[!htb]
    \centering
    \input{figures/tex_files/gan_visualization.tex}
    \caption{Approach for image colorization using conditional GANs}
    \label{fig:gan_visualization}
\end{figure*}
  Let $x \in X$ be a low resolution or a grayscale image and $y \in Y$ be it's underlying distribution from the latent space Y. Generator $G$ takes in input $x$ and produces an output $\hat{y}$. We define the mapping $x \rightarrow \hat{y}$ in the following manner:
  $$G(x) = \hat{y}$$

 The discriminative network D is fed the fabricated mapping $x \rightarrow \hat{y}$ and the underlying distribution of $x$ i.e. $y \in Y$. The network D then produces a result that is a probability distribution of the input space indicating the class of the image that it thinks the input belongs to. We define this as:
 $$D\big(G(x),y\big) = p$$
 where $p \in (0,1)$ is the probability that the image is fabricated or real.
With conditional GAN, both generator and discriminator are conditioning on the input $x$. Let the generator be parameterized by $\theta_g$ and the discriminator be parameterized by $\theta_d$. The minimax objective function can be defined as:
	\begingroup\makeatletter\def\f@size{7}\check@mathfonts \[
		\min_{\theta_g}\max_{\theta_d}\Big[\mathbb{E}_{x,y\sim p_{data}} \log D_{\theta_d}(x,y) + E_{x\sim p_{data}} \log(1 - D_{\theta_d}(x, G_{\theta_g}(x))\Big]
	\] \endgroup
    \normalsize
Where, $G_{\theta_{g}}$ is the output of the generator and $D_{\theta_d}$ is the output of the discriminator.
We're currently not introducing any noise in our generator to keep things simple for the time being. Also, we consider $L1$ difference between input $x$ and output $y$ in generator. On each iteration, the discriminator would maximize $\theta_d$ according to the above expression and generator would minimize $\theta_g$ in the following way:
	\[
		\min_{\theta_g}\Big[-\log(D_{\theta_d}(x,G_{\theta_g}(x)))+\lambda \Vert G_{\theta_g}(x) - y \Vert_1 \Big]
	\]

\subsection{Transferred learning and model tweaking}
\hspace*{0.25 in}\cite{isola2018imagetoimage} proposed a general solution to many image-to-image translation tasks. We propose a similar methodology as proposed in the paper with a few minor tweaks, so as to significantly reduce the amount of training data needed and minimize the training time to achieve similar results. Initially, we propose the use of a U-net for the generator. A pre-trained ResNet-18 network was used as the \textit{encoder} for the U-net with half of it's layers clipped off so as to get feature abstractions of the intermediate layers. \cite{ledig2017photorealistic} showed that training the generator separately in a supervised, deterministic manner helps it generalize the mapping from the input space to outputs. The idea solves the problem of \textit{"The blind leading the blind"} that is persistent in most image-to-image translation tasks to date. \\
\hspace*{0.25 in}We pre-train our generator independently using the imagenet weights over the Common Objects in Context (COCO) dataset. Unlike adversarial training, this phase was supervised and the loss function used was \textit{mean absolute error} or the \textit{L1 Norm} from the target image. Though trained deterministically,the problem of rectifying incorrect predictions still persists due to constraints over the convergence of loss metric. To combat this, this trained generator was trained once again in an adversarial fashion to help generalize it further. We hypothesize that re-training in an adversarial fashion will further rectify the subtle color differences that \textit{mae} couldn't solve.\\
\hspace*{0.25 in}A pre-trained ResNet-50 with imagenet weights was used as the discriminator with last few layers clipped off. The discriminative network used is something called a "Patch Discriminator" proposed by \cite{isola2018imagetoimage}. In a \textit{vanilla} discriminator proposed by \cite{goodfellow2014generative}, the network produces a scalar output, representing the probability that the data $x$ belongs to the input distribution and not the generator distribution $p_g$. \cite{isola2018imagetoimage} proposed a modification to the discriminative network so as to produce, instead of a single scalar, a vector of probabilities representing different localities of the input distribution (image) $x$. For instance, a discriminator producing a $50 \times 50$ vector represents the probabilities every receptive field that is covered by the output vector. Thus, we can localize the corrections in the image that the generator should make.\\
\hspace*{0.25 in}Finally, the trained generator and trained discriminator are fine-tune to fit on our data which is rather small in size compared to the previous datasets. These networks are trained in an adversarial fashion using the conditional GAN objective function \citep{isola2018imagetoimage} with some noise introduced as the L1 norm of generated image tensor to the target image tensor. The reason behind this is to train the generator using an adversarial component while trying to minimize the Manhattan distance between the generator output and target vector space.
\subsection{Experimental Setup}
\hspace*{0.25 in}We use a mini-batch gradient descent with a batch size of 10, 16 and 32 for different iterations. We use Adam optimizer with $\beta_1 = 0.5$ and $\beta_2 = 0.999$. The generator and discriminator has a learning rate of $2e-4$ which remains constant throughout the training. We train the model with different epochs ranging between 20,50 and 100, saving the best model weights determined by L1 norm between the output of the generator and the target image. We use early stopping with a patience value of 10. The image size is $64\times64$. Our implementation is using Python, numpy, Tensorflow and tf-keras. It takes about 24 hours to complete training over the COCO dataset and about 12 to 13 hours to fit the model on given astronomical data on a NVIDIA Tesla K80 GPU.

\section{Results and Discussions}
\hspace*{0.25 in}In the following section, we briefly compare and summarize the results of the implemented architectures and evaluate their performance. The evaluation is done qualitatively as well as quantitatively. We compare the performance of each model using L1 and L2 norm distance between the predictions and targets. Though unreliable, this method provides us with a somewhat decent ground to perform a comparative study and evaluate the reliability of such metrics on GAN evaluation compared to qualitative, visual evaluation. We also measure the performance of the GANs by using Fréchet inception distance (FID) as an attempt to quantify how close the produced images are to the original ones.\\
\hspace*{0.25 in}To evaluate the model performance by virtue of convergence of the objective function, we plot the generator loss throughout the training on two different networks.\\
\begin{figure}[!htb]
	\centering
	\begin{subfigure}[b]{0.2\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/resnet18_generator_loss}
		\caption{ResNet18 U-net Generator}
		\label{fig: resnet_loss}
	\end{subfigure}
		\hspace{0.1 in}
	\begin{subfigure}[b]{0.2\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/resnet18_discriminator_loss}
		\caption{ResNet50 discriminator loss}
		\label{fig: resnet50_disc_loss}
	\end{subfigure}
	\caption{Curve plots for ResNet18 U-net \& ResNet50 discriminator}
	\label{fig: resnet18_curves}
\end{figure}
Figure \ref{fig: resnet_loss} and \ref{fig: resnet50_disc_loss} shows that the generator converges quite nicely. The discriminator on the other hand oscillates because of the convergence that the generator shows. GAN losses are pretty non-intuitive but we can draw some observations from the loss curves. It seems that the pattern of oscillation and convergence repeats when networks are trained in an adversarial fashion. It is observed that the ResNet U-net with a pre-trained ResNet-18 for its backbone converges decently in the beginning but later shows spikes when nearing the end of training loop. It seems that the loss convergence doesn't signify whether the model is predicting expected results. The loss function just converges to the minimum value that the cost function descends to, permitted by the learning rate. It would normally signify that the GAN has found some optimum point in the vector space that is at the lowest potential and cant decrease any further. It essentially means that the GAN has leaned enough. Due to the large number of dimensions, owing to the high amount of trainable variables, such combinations, where the function converges, can be huge in volume. Thus, these numbers don't provide any better understanding of the bias or variance the model is facing. Also, we discover that if the loss hasn't converged well, it doesn't necessarily mean that the model hasn't learned anything. On visual inspection, the generated results show similar results to the ground truth, even with high generator losses. This might be due to presence of a content loss parameter in the loss function where we try to minimize the function by  minimizing the L1 norm between the generator predictions and target.

@@

Table \ref{tab:Super-resolution_results} shows the L1 and L2 distances of predicted results by each model with the ground truth image. It is observed that Ledig's SRGAN, after a bit of fine-tuning performs really well in comparison to the the pre-trained version. To further improve this, \cite{lim2017enhanced} proposed an optimized version of SRGAN by removing the unnecessary modules in the conventional resnets and showed that Enhanced Deep Residual Networks performed better at upscaling task. When trained in an adversarial manner, the ED-SRGAN performs better than the traditional SRGAN. \cite{yu2018wide} further improved the idea by increasing the widening factor ($\times 2$ and $\times 4$) to ($\times 6$ and $\times 9$). This Wide Activation Deep Super Resolution network further improved the performance for single image super-resolution. When we implement this in an adversarial manner, we achieve excellent results. It is evident from the results that the best performing network is the WDSR.
